\vspace{-.5cm} Lights: a generalized joint model for high-dimensional multivariate longitudinal data and censored durations \vspace{.5cm}

### 2021-09-01 paper.tex
Add the matrix form of the updated baseline hazard function

Let us introduce the following useful notations
\begin{equation*}
  \left\{
    \begin{array}{ll}
    \tilde g^1&: b_i \mapsto b_i,\\
    \tilde g^2&: b_i \mapsto b_ib_i^\top, \\
    \tilde g^3_{s}&: b_i \mapsto \Psi_i(s), \\
    \tilde g^4_{s, \gamma_k}&: b_i \mapsto \exp \big\{\gamma_k^\top g^3_{s}(b_i) \big\}, \\
    \tilde g^5_{s, \gamma_k}&: b_i \mapsto g^3_{s}(b_i)^\top g^4_{s}(b_i).
    \end{array}
  \right. 
\end{equation*}
We also denote 
\[ \cI_1 = 
\begin{bmatrix}
  \ind{\{t_{1} = t_1^u\}}  & \cdots & \ind{\{t_{1} = t_J^u\}}\\
  \vdots &  \ddots & \vdots \\
  \ind{\{t_{n} = t_1^u\}} & \cdots & \ind{\{t_{n} = t_J^u\}}
\end{bmatrix}
\in \R^{n \times J}
\quad \text{and} \quad 
\cI_2 = 
\begin{bmatrix}
  \ind{\{t_{1} \leq t_1^u\}} & \cdots & \ind{\{t_{1} \leq t_J^u\}}\\
  \vdots &  \ddots & \vdots \\
  \ind{\{t_{n} \leq t_1^u\}} & \cdots & \ind{\{t_{n} \leq t_J^u\}}
\end{bmatrix}
\in \R^{n \times J} \]
some matrices to compare each event times with the vector of distinct event times. Those matrices are widely used in practice to replace indicators. For instance
\begin{align*}
f(t_i, \delta_i| b_i, G_i = k ; \theta^{(w)}) = &\big(\lambda_0^{(w)}(t_i) g^4_{t_i, \gamma_k^{(w)}}(b_i) \big)^{\delta_i} \\ 
&\times \exp \Big\{ - \sum_{j=1}^J \lambda_0^{(w)}(t_j^u) g^4_{t_j^u, \gamma_k^{(w)}}(b_i) \ind{\{t_j^u \leq t_i\}} \Big\}
\end{align*}
so that one can compute the tensor
\[ \underset{J}{\text{sum}} \big( \boldsymbol{g^4_{\gamma_k}} \boldsymbol{\lambda_0^{(w)}} \cI_1 \big)^{\boldsymbol{\delta}} \exp \big\{- \underset{J}{\text{sum}} \big(\boldsymbol{g^4_{\gamma_k}} \boldsymbol{\lambda_0^{(w)}} \cI_2 \big) \big\} \in \R^{n \times 2N} \]
where 
\begin{equation*}
  \left\{
    \begin{array}{ll}
    \boldsymbol{g^4_{\gamma_k}} \in \R^{2N \times J} \text{ with } \big[\boldsymbol{g^4_{\gamma_k}}\big]_{i, j} = g^4_{t_j^u, \gamma_k^{(w)}}(\breve{b}_i) , \\
    \boldsymbol{\lambda_0^{(w)}} = \big(\lambda_0^{(w)}(t_1^u), \ldots, \lambda_0^{(w)}(t_J^u) \big)^\top \in \R^J, \\
    \boldsymbol{\delta} = (\delta_1, \ldots, \delta_n)^\top \in \R^n.
    \end{array}
  \right. 
\end{equation*}

>>>>

Let us introduce the following useful notations
\begin{equation*}
  \left\{
    \begin{array}{ll}
    \tilde g^1&: b_i \mapsto b_i,\\
    \tilde g^2&: b_i \mapsto b_ib_i^\top, \\
    \tilde g^3_{s}&: b_i \mapsto \Psi_i(s), \\
    \tilde g^4_{s, \gamma_k}&: b_i \mapsto \exp \big\{\gamma_k^\top g^3_{s}(b_i) \big\}, \\
    \tilde g^5_{s, \gamma_k}&: b_i \mapsto g^3_{s}(b_i)^\top g^4_{s}(b_i).
    \end{array}
  \right. 
\end{equation*}
We also denote 
\[ \cI_1 = 
\begin{bmatrix}
  \ind{\{t_{1} = t_1^u\}}  & \cdots & \ind{\{t_{1} = t_J^u\}}\\
  \vdots &  \ddots & \vdots \\
  \ind{\{t_{n} = t_1^u\}} & \cdots & \ind{\{t_{n} = t_J^u\}}
\end{bmatrix}
\in \R^{n \times J}
\quad \text{and} \quad 
\cI_2 = 
\begin{bmatrix}
  \ind{\{t_{1} \leq t_1^u\}} & \cdots & \ind{\{t_{1} \leq t_J^u\}}\\
  \vdots &  \ddots & \vdots \\
  \ind{\{t_{n} \leq t_1^u\}} & \cdots & \ind{\{t_{n} \leq t_J^u\}}
\end{bmatrix}
\in \R^{n \times J} \]
some matrices to compare each event times with the vector of distinct event times. Those matrices are widely used in practice to replace indicators. For instance
\begin{align*}
f(t_i, \delta_i| b_i, G_i = k ; \theta^{(w)}) = &\big(\lambda_0^{(w)}(t_i) g^4_{t_i, \gamma_k^{(w)}}(b_i) \big)^{\delta_i} \\ 
&\times \exp \Big\{ - \sum_{j=1}^J \lambda_0^{(w)}(t_j^u) g^4_{t_j^u, \gamma_k^{(w)}}(b_i) \ind{\{t_j^u \leq t_i\}} \Big\}
\end{align*}
so that one can compute the tensor
\[ \underset{J}{\text{sum}} \big( \boldsymbol{g^4_{\gamma_k}} \boldsymbol{\lambda_0^{(w)}} \cI_1 \big)^{\boldsymbol{\delta}} \exp \big\{- \underset{J}{\text{sum}} \big(\boldsymbol{g^4_{\gamma_k}} \boldsymbol{\lambda_0^{(w)}} \cI_2 \big) \big\} \in \R^{n \times 2N} \]
or the updated baseline hazard function~\eqref{eq:update_lambda_0} can be rewrote as
\begin{equation*}
  \label{eq:update_lambda_0_matrix}
  \lambda_0^{(w+1)}(t)= \dfrac{\underset{n}{\text{sum }} \boldsymbol{\delta} \cI_1 }{\underset{n}{\text{sum }} \underset{K = 2}{\text{sum }} \hat \E_{\theta^{(w)}}[\boldsymbol{g^4}] \cI_2 \boldsymbol{\hat \pi^{\theta^{(w)}}}  },
\end{equation*}
where 
\begin{equation*}
  \left\{
    \begin{array}{ll}
    \boldsymbol{g^4_{\gamma_k}} \in \R^{2N \times J} \text{ with } \big[\boldsymbol{g^4_{\gamma_k}}\big]_{i, j} = g^4_{t_j^u, \gamma_k^{(w)}}(\breve{b}_i) , \\
    \boldsymbol{\lambda_0^{(w)}} = \big(\lambda_0^{(w)}(t_1^u), \ldots, \lambda_0^{(w)}(t_J^u) \big)^\top \in \R^J, \\
    \boldsymbol{\delta} = (\delta_1, \ldots, \delta_n)^\top \in \R^n, \\
    \boldsymbol{\hat \pi^{\theta^{(w)}}} \in \R^{2 \times n} \text{ with }  \boldsymbol{\hat \pi_1^{\theta^{(w)}}}= (\hat \pi_{11}^{\theta^{(w)}}, \ldots, \hat \pi_{n1}^{\theta^{(w)}})^\top \in \R^n \text{ and } \boldsymbol{\hat\pi_0^{\theta^{(w)}}} = 1 - \boldsymbol{\hat\pi_1^{\theta^{(w)}}}.
    \end{array}
  \right. 
\end{equation*}
